From Reversible Programs to \\ Univalent Universes and Back

### 2021-07-09 Pi+/paper/discussion.tex
more on sec 8

@@

%%%%%%%%%%%%%%
\paragraph{Algebraic Theories} In universal algebra, algebraic theories are used to describe algebraic structures, such
as groups or rings. A specific group or ring is a model of the appropriate algebraic theory. Algebraic theories are
usually \emph{presented} in terms of logical syntax, that is, as first-order theories whose signatures allow only
functional symbols, and whose axioms are universally quantified equations. In his seminal
thesis~\cite{lawvereFUNCTORIALSEMANTICSALGEBRAIC1963}, Lawvere defined a presentation-free categorical notion of
universal algebraic structure, called a Lawvere theory.

Programming Languages, such as the $\lambda$-calculus, can be viewed as algebraic structures with variable-binding
operators, which can be formalised using second-order algebraic theories~\cite{fioreSecondOrderAlgebraicTheories2010},
or algebraic theories with closed structure~\cite{hylandClassicalLambdaCalculus2017}, called $\lambda$-theories, making
the $\lambda$-calculus the presentation of the initial $\lambda$-theory $\Lambda$.

Our family of reversible languages have been presented as first-order algebraic
2-theories~\cite{cohenCoherenceRewriting2theories2009,bekeCategorificationTermRewriting2011,yanofskySyntaxCoherence2000},
which are a categorification of algebraic theories. The types $\zerot$ and $\onet$ are nullary function symbols, the
type formers $+$ and $\times$ are binary function symbols, the 1-combinators are invertible reduction rules, and the
2-combinators are equations or coherence diagrams of compositions of reduction rules. Just like models of Lawvere
theories are given by algebras of (finitary) monads on $\SetCat$, models of 2-theories are given by algebras of 2-monads
on $\CatCat$. The particular one we're interested in here is the free symmetric monoidal completion 2-monad.

>>>>

%%%%%%%%%%%%%%
\paragraph{Algebraic Theories} In universal algebra, algebraic theories are used to describe structures such as groups
or rings. A specific group or ring is a model of the appropriate algebraic theory. Algebraic theories are usually
\emph{presented} in terms of logical syntax, that is, as first-order theories whose signatures allow only functional
symbols, and whose axioms are universally quantified equations. In his seminal thesis,
\citet{lawvereFUNCTORIALSEMANTICSALGEBRAIC1963} defined a presentation-free categorical notion of universal algebraic
structure, called a Lawvere theory. Programming Languages, such as the $\lambda$-calculus, can be viewed as algebraic
structures with variable-binding operators, which can be formalised using second-order algebraic
theories~\cite{fioreSecondOrderAlgebraicTheories2010}, or algebraic theories with closed
structure~\cite{hylandClassicalLambdaCalculus2017}, called $\lambda$-theories, making the $\lambda$-calculus the
presentation of the initial $\lambda$-theory $\Lambda$. Our family of reversible languages $\PiLang$ have been presented
as first-order algebraic
2-theories~\cite{cohenCoherenceRewriting2theories2009,bekeCategorificationTermRewriting2011,yanofskySyntaxCoherence2000},
which are a categorification of algebraic theories. The types $\zerot$ and $\onet$ are nullary function symbols, the
type formers $+$ and $\times$ are binary function symbols, the 1-combinators are invertible reduction rules, and the
2-combinators are equations or coherence diagrams of compositions of reduction rules. Just like models of Lawvere
theories are given by algebras of (finitary) monads on $\SetCat$, models of 2-theories are given by algebras of 2-monads
on $\CatCat$. The particular one related to our development is the free symmetric monoidal completion 2-monad.

@@
